Automated Example Oriented REST API Documentation

### 2017-08-16 introduction.tex
spell corrections

REST APIs are a type of APIs that are used to integrate software components using web technologies. Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}. For example, GitHub, an online code repostory hosting tool, has a REST API that can be used by third-party applications to integrate with GitHub features. The GitHub REST API\footnote{\url{https://developer.github.com/v3/repos/\#create}} has a resource called \texttt{Repository} to denote a code repository that can be identified by the URL \url{/user/repos}. To create a new \texttt{Repository} for a user, the GitHub API can be invoked via HTTP $POST$ at \url{/user/repos} using the following JSON representation of a \texttt{Repository}:

@@

In today's world of technology, REST APIs have become ubiquitous and the primary choice for integrating Internet enabled applications due to its simplicity and similarity with HTTP \cite{mangler2010origin}. For example, a real estate listing website uses a REST API to collect ``walk score'' and another REST API to show a map view of a property listing. By incorporating the map and walk score using REST APIs, the real estate listng site provides important information to their users. Most REST APIs, including these two examples, are often documented manually or using custom implemented tools specific to the APIs that are not publicly available. This requires effort to generate and maintain the documentation of REST APIs over time since there is a lack of reusable tool support.

@@

While there are overlaps between the documentation requirements of local and REST APIs, there are significant requirements that are unique to each. The documentation of most local APIs contain a structrual definition of the API classses and methods. API usage examples for local APIs are documented using the same programming language of the API. On the other hand, REST APIs are defined using HTTP specific parameters such as API endpoints, query parameters, HTTP headers and body. REST API clients can use any programming language agnostic to the API implementation language. Due to such structural and implemenation differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.

@@

\begin{itemize}
  \item \textit{RG1: To identify a set of REST API documentation requirements.} Existing research on API usability has derived a number of API documentation requirements primarily focusing on local APIs. Due to the structural and implementation differences between local and REST APIs, I aimed to identity REST API specific documentation requirements in my research. This has several implicatons as follows: having access to a scientifically derived list of REST API documentation requirements allows REST API developers to make an informed decision related to planning and execution of their documentation efforts. The requirements provide a background research that can be used by practitioners to compare and contrast existing REST API documentation tools while selecting an appropriate tool for their API. REST API documentation tool developers can also use the requirements to build necessary features.
  \item \textit{RG2: To develop and evaluate a technique to overcome the commonly observed but unresolved challenges related to REST API documentation.} I aimed to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods. It has the following implications: REST API developers can utilize the technique to overcome the common challenges with the generation and maintenance of their API documentation. REST API client developers can access a usable API documentation.
\end{itemize}

@@

\begin{itemize}
  \item \textit{RO1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the proposed solutions from existing literature. This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a neccessary background work for improving the state of REST API documentation process.
  \item \textit{RO3: What technique can be used to document REST APIs to meet the requirements?} This is the central research objective. I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
  \item \textit{RO4: Is the proposed technique feasible for REST API developers?} An evaliation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
  \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developers.
\end{itemize}

\section{Cohesion and Unitary Report}
To address the aformentioned research goals and objectives, my research strategy included five sub-projects, result of which have been published as papers in peer-reviewed international venues. Table~\ref{table:chapter_summary} shows a high level overview of the flow of arguments between the sub-projects to meet the research goals.

\begin{table}
  \caption{Summary and Flow of Arguments of the Sub-projects}
  \begin{tabular}{|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{4.5cm}|p{6cm}|}
  \hline
  Chapter & Goal & Objective & Research Project & Primary Outcome \\
  \hline
  Chapter~\ref{chapter:case_study}  & RG1 & RO1 & A case study of Web API evolution \cite{sohan2015case}. & A set of challenging problems related to evolving Web APIs. Selected REST API documentation as the problem of interest for this research. \\
  \hline
  Chapter~\ref{chapter:spy_rest} & RG1 and RG2 & RO2 and RO3 & SpyREST: Automated RESTful API documentation using an HTTP proxy server \cite{sohan2015spyrest}. & A list of requirements and the concept of using interception to generate REST API documentation to satisfy the requirements. \\
  \hline
  Chapter~\ref{chapter:demo_paper} & RG2 & RO3 & SpyREST in action: An automated RESTful API documentation tool \cite{sohan2015spyrest_tool}. & A reusable tool and a demonstration of the proposed REST API documentation technique. \\
  \hline
  Chapter~\ref{chapter:cisco_study} & RG2 & RO4 & Automated example oriented REST API documentation at cisco \cite{sohan_cisco}. & An empirical evidence that the proposed technique is feasible for practical use, and a list of problems for potential future research.\\
  \hline
  Chapter~\ref{chapter:controlled_study} & RG2 & RO5 & A study of the effectiveness of usage examples in REST API documentation \cite{sohan_vlhcc}. & An empirical evidence that REST API client developers can overcome common obstacles when the documentation contains usage examples, as suggested in the proposed technique.\\
  \hline
\end{tabular}
\label{table:chapter_summary}
\end{table}

The first research project is described in Chapter~\ref{chapter:case_study}, where I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs \cite{sohan2015case}. The results are based on a study of nine industrical APIs. I found a lack of reusable techniques in the existing literature to automatically generate and maintain the documentation of multi-version evolving REST APIs with usage examples. I persued further research to solve this problem as discussed in the following paragraphs.

In Chapter~\ref{chapter:spy_rest}, I discuss my second research project where I derived a set of REST API documentation requirements and proposed a new technique to satisfy the requirements \cite{sohan2015spyrest}. Here, I present a list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}. I discuss the available REST API documentation techniques and compare against the requirements. I present, SpyREST, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and autmatically synthesize API traffic to generate and update API documentation to satisfy the identified requirements. The primary advantage of the presented technique is that the documentation is generated from executable code. As a result, a continuously updated REST API documentation with realistic usage examples is achievable with a reduced manual effort.

I present my third research project, related to RG2, in Chapter~\ref{chapter:demo_paper}, where I present a prototype implementation of the aformentioned technique \cite{sohan2015spyrest_tool}. In this prototype implementation, I discuss the process that transforms a number of example HTTP requests and responses into a usable REST API documentation. I also present an example where automated test code is used to generate and maintain a continuously updated REST API documentation using the developed tool.

@@

I present my fifth research project, related to RG2, in Chapter~\ref{chapter:controlled_study}, where I discuss a controlled study to undestand the impact of usage examples in API documentation on REST API client developers \cite{sohan_vlhcc}. I observed that providing usage examples with data taken from API test code reduces the obstacles and improves API client developer success rate, reduces required trial attempts, and improves satisfaction ratings. The findings from this study provides an empirical evidence that by using the proposed technique of generating REST API documentation from API test code can help API client developers to achive a higher level of success with a reduced effort.

In Chatper~\ref{chapter:conclusion}, I conclude by discussing how the research objectives are achieved and provide a list of problems for future research.

@@

\begin{enumerate}
  \item A list of REST API documentation requirements derived from the existing literature and the current state of industry practices to be used as a guideline by practitioners and researchers.
  \item SpyREST, a novel technique and a reusable tool that uses interception to automatically generate and maintain evolving REST API documentation with usage examples from executable code.
  \item An industrial evaluation of SpyREST to demonstrate the feasibility of the proposed technique that practitioners can follow to improve their REST API documentation experience.
  \item An empricial evidence showing the common obstacles faced by REST API client developers that can be reduced with usage examples to help the prioritization of REST API documentation efforts.
\end{enumerate}

>>>>

REST APIs are a type of APIs that are used to integrate software components using web technologies. Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}. For example, GitHub, an online code repository hosting tool, has a REST API that can be used by third-party applications to integrate with GitHub features. The GitHub REST API\footnote{\url{https://developer.github.com/v3/repos/\#create}} has a resource called \texttt{Repository} to denote a code repository that can be identified by the URL \url{/user/repos}. To create a new \texttt{Repository} for a user, the GitHub API can be invoked via HTTP $POST$ at \url{/user/repos} using the following JSON representation of a \texttt{Repository}:

@@

In today's world of technology, REST APIs have become ubiquitous and the primary choice for integrating Internet enabled applications due to its simplicity and similarity with HTTP \cite{mangler2010origin}. For example, a real estate listing website uses a REST API to collect ``walk score'' and another REST API to show a map view of a property listing. By incorporating the map and walk score using REST APIs, the real estate listing site provides important information to their users. Most REST APIs, including these two examples, are often documented manually or using custom implemented tools specific to the APIs that are not publicly available. This requires effort to generate and maintain the documentation of REST APIs over time since there is a lack of reusable tool support.

@@

While there are overlaps between the documentation requirements of local and REST APIs, there are significant requirements that are unique to each. The documentation of most local APIs contain a structural definition of the API classes and methods. API usage examples for local APIs are documented using the same programming language of the API. On the other hand, REST APIs are defined using HTTP specific parameters such as API endpoints, query parameters, HTTP headers and body. REST API clients can use any programming language agnostic to the API implementation language. Due to such structural and implementation differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.

@@

\begin{itemize}
  \item \textit{RG1: To identify a set of REST API documentation requirements.} Existing research on API usability has derived a number of API documentation requirements primarily focusing on local APIs. Due to the structural and implementation differences between local and REST APIs, I aimed to identity REST API specific documentation requirements in my research. This has several implications as follows: having access to a scientifically derived list of REST API documentation requirements allows REST API developers to make an informed decision related to planning and execution of their documentation efforts. The requirements provide a background research that can be used by practitioners to compare and contrast existing REST API documentation tools while selecting an appropriate tool for their API. REST API documentation tool developers can also use the requirements to build necessary features.
  \item \textit{RG2: To develop and evaluate a technique to overcome the commonly observed but unresolved challenges related to REST API documentation.} I aimed to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods. It has the following implications: REST API developers can utilize the technique to overcome the common challenges with the generation and maintenance of their API documentation. REST API client developers can access a usable API documentation.
\end{itemize}

@@

\begin{itemize}
  \item \textit{RO1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the proposed solutions from existing literature. This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a necessary background work for improving the state of REST API documentation process.
  \item \textit{RO3: What technique can be used to document REST APIs to meet the requirements?} This is the central research objective. I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
  \item \textit{RO4: Is the proposed technique feasible for REST API developers?} An evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
  \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developers.
\end{itemize}

\section{Cohesion and Unitary Report}
To address the aforementioned research goals and objectives, my research strategy included five sub-projects, result of which have been published as papers in peer-reviewed international venues. Table~\ref{table:chapter_summary} shows a high level overview of the flow of arguments between the sub-projects to meet the research goals.

\begin{table}
  \caption{Summary and Flow of Arguments of the Sub-projects}
  \begin{tabular}{|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{4.5cm}|p{6cm}|}
  \hline
  Chapter & Goal & Objective & Research Project & Primary Outcome \\
  \hline
  Chapter~\ref{chapter:case_study}  & RG1 & RO1 & A case study of Web API evolution \cite{sohan2015case}. & A set of challenging problems related to evolving Web APIs. Selected REST API documentation as the problem of interest for this research. \\
  \hline
  Chapter~\ref{chapter:spy_rest} & RG1 and RG2 & RO2 and RO3 & SpyREST: Automated RESTful API documentation using an HTTP proxy server \cite{sohan2015spyrest}. & A list of requirements and the concept of using interception to generate REST API documentation to satisfy the requirements. \\
  \hline
  Chapter~\ref{chapter:demo_paper} & RG2 & RO3 & SpyREST in action: An automated RESTful API documentation tool \cite{sohan2015spyrest_tool}. & A reusable tool and a demonstration of the proposed REST API documentation technique. \\
  \hline
  Chapter~\ref{chapter:cisco_study} & RG2 & RO4 & Automated example oriented REST API documentation at Cisco \cite{sohan_cisco}. & An empirical evidence that the proposed technique is feasible for practical use, and a list of problems for potential future research.\\
  \hline
  Chapter~\ref{chapter:controlled_study} & RG2 & RO5 & A study of the effectiveness of usage examples in REST API documentation \cite{sohan_vlhcc}. & An empirical evidence that REST API client developers can overcome common obstacles when the documentation contains usage examples, as suggested in the proposed technique.\\
  \hline
\end{tabular}
\label{table:chapter_summary}
\end{table}

The first research project is described in Chapter~\ref{chapter:case_study}, where I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs \cite{sohan2015case}. The results are based on a study of nine industrial APIs. I found a lack of reusable techniques in the existing literature to automatically generate and maintain the documentation of multi-version evolving REST APIs with usage examples. I pursued further research to solve this problem as discussed in the following paragraphs.

In Chapter~\ref{chapter:spy_rest}, I discuss my second research project where I derived a set of REST API documentation requirements and proposed a new technique to satisfy the requirements \cite{sohan2015spyrest}. Here, I present a list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}. I discuss the available REST API documentation techniques and compare against the requirements. I present, SpyREST, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and automatically synthesize API traffic to generate and update API documentation to satisfy the identified requirements. The primary advantage of the presented technique is that the documentation is generated from executable code. As a result, a continuously updated REST API documentation with realistic usage examples is achievable with a reduced manual effort.

I present my third research project, related to RG2, in Chapter~\ref{chapter:demo_paper}, where I present a prototype implementation of the aforementioned technique \cite{sohan2015spyrest_tool}. In this prototype implementation, I discuss the process that transforms a number of example HTTP requests and responses into a usable REST API documentation. I also present an example where automated test code is used to generate and maintain a continuously updated REST API documentation using the developed tool.

@@

I present my fifth research project, related to RG2, in Chapter~\ref{chapter:controlled_study}, where I discuss a controlled study to understand the impact of usage examples in API documentation on REST API client developers \cite{sohan_vlhcc}. I observed that providing usage examples with data taken from API test code reduces the obstacles and improves API client developer success rate, reduces required trial attempts, and improves satisfaction ratings. The findings from this study provides an empirical evidence that by using the proposed technique of generating REST API documentation from API test code can help API client developers to achieve a higher level of success with a reduced effort.

In Chapter~\ref{chapter:conclusion}, I conclude by discussing how the research objectives are achieved and provide a list of problems for future research.

@@

\begin{enumerate}
  \item A list of REST API documentation requirements derived from the existing literature and the current state of industry practices to be used as a guideline by practitioners and researchers.
  \item SpyREST, a novel technique and a reusable tool that uses interception to automatically generate and maintain evolving REST API documentation with usage examples from executable code.
  \item An industrial evaluation of SpyREST to demonstrate the feasibility of the proposed technique that practitioners can follow to improve their REST API documentation experience.
  \item An empirical evidence showing the common obstacles faced by REST API client developers that can be reduced with usage examples to help the prioritization of REST API documentation efforts.
\end{enumerate}

### 2017-08-16 root.tex
spell corrections

In this thesis, I present and evaluate of a novel technique to solve REST API documentaiton requirements. I present a set of REST API documentation requirements by studying the existing literature and the current industry practice. From this study, I observed that REST APIs evolve frequently, but used a manual or bespoke approach for generating and maintaining their documentation. I present a reusable technique with a prototype implementation to automate the REST API documentation process with executable API usage examples. The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation. I present an industrial evaluation of the proposed technique based on a period of eighteen months of production use. From this study, I found that it was feasible to leverage API test code to automatically generate an always-updated REST API documentation for evolving REST APIs with usage examples. To evaluate the impact of usage examples on REST API client developers, I performed a controlled study with experienced REST API developers. From this case study, I found that REST API client developers faced patterns of obstacles using REST APIs that can be reduced by including API usage examples in the API documentation, as suggested in the proposed technique.

@@

In this thesis, I presented a set of five primary research objectives related to the topic of REST API documentation. I have addressed the first research objective, RO1, of finding the gap between the current industry practices and existing research on REST API documentation by performing a case study in my first research sub-project. I started by reviewing the existing literature on Web API evolution. From the existing empirical research on Web API evolution, I found that researchers have focused on profiling SOAP based web services to identify the common change patterns. They suggested focusing on the maintenance efforts related to Web API evolution because of its lingering effects. In the existing literature, I found several techniques proposed by researchers related to supporting the evolution of SOAP APIs. Analyzing the existing literature on tool support for Web API evolution, I found tools developed by researchers that primarly rely on custom formatted API specification documents that require a manual effort to generate and maintain. In this sub-project, I have also analyzed the evolution of nine Web APIs in production, selected based on their industry, popularity, maturity, and source code access to understand the current state of practice around Web API versioning, documentation, and change comminucation patterns. By comparing and contrasting the findings from the literature review and the industry practices, I observed that it was important to carry out further research to find techniques to support versioning, documentation, and change communication needs for evolving REST APIs.

I addressed RO2, finding RESAPI T documentation requirements, in my second research sub-project. I started by analyzing the existing literature on API usability to summarize a list of general API documentation requirements as derived by other researchers. Then, I analyzed the existing literature on REST API documentation. Based on these two analyses, I derived the following list of documentation requirements specific for evolving REST APIs: automated, example based, executable, version aware, customizable, reusable, and collaborative. These requirements are used as a background work to address my subsequent research objectives.

RO3, the primary objective of finding a new technique to satisfy the documentation requirements of evolving REST APIs, is addressed in my third and fourth sub-projects. I observed that most existing techniques for REST API documentation relied on custom API specification formats as an input. This approach requires manual effort to generate and maintain the API specification as a separate artifact with the API as the API evolves and fails to satisfy the identified REST API documentation requirements. I proposed a novel technique called SpyREST to automatically intercept and transform example REST API calls into usable API documentation following the identified requirements. An HTTP proxy server is used to collect data from example API calls. The collected data is then synthesized to automatically infer the structural elements of the API such as the version, API resource, HTTP verb, API action, query parameters, request and response body, HTTP headers. In addition to the structure, executable API examples are automatically constructed with realistic data. I developed a prototype tool to demonstrate a concrete implementation of the proposed concept. In a preliminary case study with three evolving REST APIs, I have shown that it is possible to use the proposed technique to generate more a accurate documentation of REST APIs compared to their official documentation that are  manually maintained.

I addressed RO4, an evaluation of the proposed technique from the perspective of REST API developers, in my fourth sub-project using a case study. I presented the proposed technique and the prototype tool to a team of software engineers at Cisco when we were in the process of selecting a documentation method for a suit of new REST APIs to be built. I discussed the selection process behind choosing SpyREST as documentation tool for this project. In this case study, I discussed how the proposed technique of using test code to drive REST API documentation has been leveraged in Cisco to create and maintain an always-updated and accurate documentation for an evolving REST API in production for a period of eighteen months. I gathered feedback from the stakeholders including the API developers, QA engineers, project managers, and customer support engineers to evaluate the merits and shortcomings of the proposed techqniue based on their production use. The evaluation results show that it is feasible to use API test code to generate and maintain REST API documentation for an evolving API with the identified REST API documentation requirements. From this evaluation, I also found that different stakeholders are able to establish a fast and continuous feedback loop about the API and its documentation while developing the API features when the API documentation is auto updated from executable code. Additionally, from this evaluation, I found a list of limitations of the proposed technique related to the lack of support for using version control features for custom edited API documentation content and automatically generated changelogs showing the differences between API elements of different API versions. Future research can be carried out to address these limitations.

My last research objective, RO5, to evaluate the impact of the proposed technique on REST API client developers, was addressed in my fifth sub-project. I performed a controlled study with 26 professional software engineers with prior REST API experience. The participants were asked to perform tasks using a REST API following its accopmpanying documentation. I enhanced an existing API documentation with usage examples leveraging the API's test code. This enhancement mimics the outcome of the proposed technique but does not use the proposed tool. Participants were mapped into two groups. One of the groups used the original API documentation and the other group used the enhanced documentation. I provided a web based API explorer so that participants could only focus on the required API inputs instead of writing a full API client application to perform the tasks. I observed that participants using the original API documentation faced similar patterns of obstacles involving the use of correct data types, data formats, HTTP headers, and request body. Participants using the enhanced documentation were able to reduce these obstacles to peform API tasks with a higher success rate, requiring less trial API calls and effort, and a higher satisfaction rating. This provides empirical evidence that the proposed technique has a positive impact on the REST API client developers. I also derived a list of recommentations related to the content of usage examples for REST API documentation that can be used by practitioners as a guideline to prioritize their documentation efforts. From this study, I also observed that participants used more trial API calls for API tasks that involve multiple dependent API calls even while using the enhanced documentation. Future research is needed to generate API documentation with examples for complex API usage scenarios to reduce this obstacle.

>>>>

In this thesis, I present and evaluate of a novel technique to solve REST API documentation requirements. I present a set of REST API documentation requirements by studying the existing literature and the current industry practice. From this study, I observed that REST APIs evolve frequently, but used a manual or bespoke approach for generating and maintaining their documentation. I present a reusable technique with a prototype implementation to automate the REST API documentation process with executable API usage examples. The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation. I present an industrial evaluation of the proposed technique based on a period of eighteen months of production use. From this study, I found that it was feasible to leverage API test code to automatically generate an always-updated REST API documentation for evolving REST APIs with usage examples. To evaluate the impact of usage examples on REST API client developers, I performed a controlled study with experienced REST API developers. From this case study, I found that REST API client developers faced patterns of obstacles using REST APIs that can be reduced by including API usage examples in the API documentation, as suggested in the proposed technique.

@@

In this thesis, I presented a set of five primary research objectives related to the topic of REST API documentation. I have addressed the first research objective, RO1, of finding the gap between the current industry practices and existing research on REST API documentation by performing a case study in my first research sub-project. I started by reviewing the existing literature on Web API evolution. From the existing empirical research on Web API evolution, I found that researchers have focused on profiling SOAP based web services to identify the common change patterns. They suggested focusing on the maintenance efforts related to Web API evolution because of its lingering effects. In the existing literature, I found several techniques proposed by researchers related to supporting the evolution of SOAP APIs. Analyzing the existing literature on tool support for Web API evolution, I found tools developed by researchers that primarily rely on custom formatted API specification documents that require a manual effort to generate and maintain. In this sub-project, I have also analyzed the evolution of nine Web APIs in production, selected based on their industry, popularity, maturity, and source code access, to understand the current state of practice around Web API versioning, documentation, and change communication patterns. By comparing and contrasting the findings from the literature review and the industry practices, I observed that it was important to carry out further research to find techniques to support versioning, documentation, and change communication needs for evolving REST APIs.

I addressed RO2, finding REST API documentation requirements, in my second research sub-project. I started by analyzing the existing literature on API usability to summarize a list of general API documentation requirements as derived by other researchers. Then, I analyzed the existing literature on REST API documentation. Based on these two analyses, I derived the following list of documentation requirements specific for evolving REST APIs: automated, example based, executable, version aware, customizable, reusable, and collaborative. These requirements are used as a background work to address my subsequent research objectives.

RO3, the primary objective of finding a new technique to satisfy the documentation requirements of evolving REST APIs, is addressed in my third and fourth sub-projects. I observed that most existing techniques for REST API documentation relied on custom API specification formats as an input. This approach requires manual effort to generate and maintain the API specification as a separate artifact with the API as the API evolves and fails to satisfy the identified REST API documentation requirements. I proposed a novel technique called SpyREST to automatically intercept and transform example REST API calls into usable API documentation following the identified requirements. An HTTP proxy server is used to collect data from example API calls. The collected data is then synthesized to automatically infer the structural elements of the API such as the version, API resource, HTTP verb, API action, query parameters, request and response body, and HTTP headers. In addition to the structure, executable API examples are automatically re-constructed with realistic data from the recorded API calls. I developed a prototype tool to demonstrate a concrete implementation of the proposed concept. In a preliminary case study with three evolving REST APIs, I have shown that it is possible to use the proposed technique to generate a more accurate documentation of REST APIs compared to their official documentation that are  manually maintained.

I addressed RO4, an evaluation of the proposed technique from the perspective of REST API developers, in my fourth sub-project using an industrial case study. I presented the proposed technique and the prototype tool to a team of software engineers at Cisco when we were in the process of selecting a documentation method for a suite of new REST APIs to be built. I discussed the selection process behind choosing SpyREST as documentation tool for this project. In this case study, I discussed how the proposed technique of using test code to drive REST API documentation has been leveraged in Cisco to create and maintain an always-updated and accurate documentation for an evolving REST API in production for a period of eighteen months. I gathered feedback from the stakeholders including the API developers, QA engineers, project managers, and customer support engineers to evaluate the merits and shortcomings of the proposed technique based on their production use. The evaluation results show that it is feasible to use API test code to generate and maintain REST API documentation for an evolving API with the identified REST API documentation requirements. From this evaluation, I also found that different stakeholders are able to establish a fast and continuous feedback loop about the API and its documentation while developing the API features when the API documentation is auto updated from executable code. Additionally, from this evaluation, I found a list of limitations of the proposed technique related to the lack of support for using version control features for custom edited API documentation content and automatically generated changelogs showing the differences between API elements of different API versions. Future research can be carried out to address these limitations.

My last research objective, RO5, to evaluate the impact of the proposed technique on REST API client developers, was addressed in my fifth sub-project. I performed a controlled study with 26 professional software engineers with prior REST API experience. The participants were asked to perform tasks using a REST API following its accompanying documentation. I enhanced an existing API documentation with usage examples leveraging the API's test code. This enhancement mimics the outcome of the proposed technique but does not use the proposed tool. Participants were mapped into two groups. One of the groups used the original API documentation and the other group used the enhanced documentation. I provided a web based API explorer so that participants could only focus on the required API inputs instead of writing a full API client application to perform the tasks. I observed that participants using the original API documentation faced similar patterns of obstacles involving the use of correct data types, data formats, HTTP headers, and request body. Participants using the enhanced documentation were able to reduce these obstacles to perform API tasks with a higher success rate, requiring less trial API calls and effort, and a higher satisfaction rating. This provides empirical evidence that the proposed technique has a positive impact on the REST API client developers. I also derived a list of recommendations related to the content of usage examples for REST API documentation that can be used by practitioners as a guideline to prioritize their documentation efforts. From this study, I also observed that participants used more trial API calls for API tasks that involve multiple dependent API calls even while using the enhanced documentation. Future research is needed to generate API documentation with examples for complex API usage scenarios to reduce this obstacle.
